---
title: "Hopkins - Introduction to R"
author: "Scott Bailey"
date: "7/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Follow instructions at https://cengel.github.io/R-intro/index.html#setup-instructions to set up R and RStudio.

## Data Analysis with R

```{r}
library(tidyverse)
library(readxl)
```

```{r}
phylum_data <- read_excel("Cobble2012.xlsx", sheet = "Phylum")

head(phylum_data, n=10)
```

```{r}
summary(phylum_data)
```

What data are we getting from this summary?

```{r}
glimpse(phylum_data)
```

We can see that the data was read with the correct data types here. The taxonomy column is a character vector, while the rest of the columns are doubles, or double precision floating point numbers. If we wanted to treat these numbers specifically as integers, we could do that, but R is actually pretty good at handling converstion from floating point to integer as needed. Since the taxa are finite and unique, we could treat those as factors, or categorical variables, but we can worry about that if we need to do an analysis where it's important.

One of the first things that we can do to start understanding data is to start graphing it. Why don't we see if we can graph the occurences of a given taxon across the different sites. We'll start by filtering our data to just the row we're interested in. When we do that, we'll interest a new operator, the `pipe` (`%>%`), as well as one of the verbs we can use with our data by way of tidyverse, `filter`.

```{r}
filter(phylum_data, taxonomy == "Annelida")
```

This is a fine way to write this code, but much of the R community has switched to using the pipe operator, so let's rewrite.

```{r}
phylum_data %>%
  filter(taxonomy == "Annelida")
```

It might help to transpose the data so that we have many columns but only one row. Instead of using a traditional transposition, we're going to use the `gather` verb to convert from a wide format to a long format. We'll also save this data to a variable so it's easier to use.

```{r}
annelida <- phylum_data %>%
  filter(taxonomy == "Annelida") %>% 
  gather("location", "value", -taxonomy)

annelida
```

With our data in the shape we'd like it, we can now send that data to ggplot to graph. Pay attention to the switch from the pipe operator connecting lines to the plus sign.

```{r}
annelida %>%
  ggplot(aes(x=location, y=value)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = -45))
```

Let's do one more thing with this to make it a bit more informatiave. Let's add a horizontal mean line so that we can tell which locations and depths have more or less than the average. 

```{r}
mean_annelida <- mean(annelida$value)

annelida %>%
  ggplot(aes(x=location, y=value)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = -45)) +
  geom_hline(aes(yintercept = mean_annelida), color = "red")
```

Instead of thinking about this as just how to add a horizontal line, think about this as how to add an arbitrary line to your graph. It could be at any valuable that's meaningful in your presentation. 

We'll worry about the colors, labels, and such later on. For now it's enough to see that we can pretty easily create graphs to start exploring the data and communicating something about it. 


Let's do one more similar task for practice. Let's suppose that we want to graph all the taxa at a specific location and depth. What do you think our steps are? Do we need to reshape the data this time?

```{r}
phylum_data %>%
  select(taxonomy, HMS_10) %>%
  ggplot(aes(x = taxonomy, y = HMS_10)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = -90))
```




Let's step back for a minute to our original data. What if we only cared about a certain location, say HMS. Let's restrict the data we look at to just that location. We'll use a new verb: select. 

```{r}
hms_data <- phylum_data %>%
  select(taxonomy, HMS_10, HMS_20, HMS_30)

hms_data
```

We can explicitly name each column we want or don't want, or we could use some helper functions. 

```{r}
hms_data <- phylum_data %>%
  select(taxonomy, starts_with("HMS"))

hms_data
```

What if we want to then add all the values for all the depths and put that data into a new column? For that, we'll use the mutate verb. 

```{r}
hms_data <- hms_data %>%
  mutate("HMS_total" = HMS_10 + HMS_20 + HMS_30)

hms_data
```

You can use mutate to create a new column based on other columns or other data, or to modify a column in place.

What if we now want to know which taxon had the largest number of occurences across all depths?

```{r}
hms_data %>%
  arrange(desc(HMS_total))
```

We've learned most of the verbs that we get with dplyr, the data manipulation part of tidyverse. Combining these verbs together you can do quite a lot with your data. Let's return to our full data again, and think about other questions we can ask of it. 

Just to remind us of what it looks like:

```{r}
head(phylum_data, n=10)
```

What if we just wanted to visualize this as is? How could we graphically represent these observations?

What if we tried a heat map? We'll need to reshape our data. We need our data such that each observation has a column for the site, the taxon, and the value. 

```{r}
phylum_long <- phylum_data %>%
  gather("location", "value", HMS_10:Pt.Pinos_40)

phylum_long %>% 
  ggplot(aes(x = location, y = taxonomy, fill = value)) +
  geom_tile() +
  geom_text(aes(label = value), color = "white", size = 2) +
  theme(axis.text.x = element_text(angle = -90)) +
  ggtitle("Heatmap of taxon observations")
```

Quick example of faceting. 

```{r}
phylum_data %>%
  gather("location", "value", -taxonomy) %>%
  ggplot(aes(x = taxonomy, y = value)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = -90)) +
  facet_wrap(~location)

```




TODO: try to determine whether the depth or the location seems most important in grouping observations.

```{r}
cor(phylum_data$HMS_10, phylum_data$HMS_20, method = "pearson")
cor(phylum_data$HMS_10, phylum_data$LovP_30)

phylum_data %>%
  select(-taxonomy) %>%
  cor()
```


